Open Questions for Computer Science and Cheminformatics

### 2011-11-21 OpenQuestions.tex
Added some more comments and fixed typos.

\item \emph{Inverse QSAR}

(statistical) QSAR is an example of traditional data mining where one
correlates a set of independent variables to a dependent
variable. This lets one predict properties of a new molecule based on
its structural features and those of a training set. The inverse QSAR
problem is such that given a molecule structure representation
(usually a descriptor vector), what are the possible input structures
that satisfy the representation. This can be made more complex, by
asking that the input structures also satisfy aexperimental property
range.
\begin{itemize}
\item What methodologies can be devised to address this problem?
\item How does a given methodology allow us to translate the
  descriptor space to a minimal region of the input space?
\item Are there internal features of a modeling algorithm (say
  hyperplanes in a SVM approach) that lets us simplify the problem?
\end{itemize}

@@

  Binary feature vectors are a common practice for chemical similarity
  searches. The typical process starts with 1. creating binary
  substructure (or other feature vectors) \cite{citeulike:8530538},
  2. creating fixed-length binary vectors of typically 1024 bits for
  reducing space requirements and speeding up further similarity
  searches (by loosing some accuracy), 3. creating further
  pre-computations for speeding up threshold based similarity searches
  \cite{doi:10.1021/ci800076s}.\\  Still, if we are interested to employ
  dynamic changes in the similarity encodings, e.g. using only a set
  of binary features, then previously done hashing or pre-computations
  might need to be redone efficiently on the fly. Finally, the major
  goal is to employ similarity searches \cite{doi:10.1021/ci200235e}
  on a scale of multiple million entries and more optimizations and
  benchmarking studies are urgently required, e.g. using GPUs
  \cite{doi:10.1021/ci1004948}, optimizing pair-wise similarity
  calculations \cite{MINF:MINF201100050}.
%

@@

  Vectorial molecule encodings can serve as efficient approximations
  of molecules.  Sometimes non-vectorial molecular 3D shape or
  molecule kernel comparisons might be more suitabe to compare
  molecules, since they might better correlate with activities. One
  key problem is that non-vectorial encodings require to compare all
  molecules (or their 3D conformational eplosions) in a pair-wise
  manner.  This becomes prohibitively expensive when considering
  millions of molecules.  Can dyadic data approaches help
  \cite{Hochreiter:2006:SVM:1159508.1159516}? Other approximations or
  cascading flows?
%
\item \emph{Using multiple annotations for improving molecular mining/predictions (chemogenomics)}

>>>>

\item \emph{Inverse QSAR (or de-novo design)}

(statistical) QSAR is an example of traditional data mining where one
correlates a set of independent variables to a dependent
variable. This lets one predict properties of a new molecule based on
its structural features and those of a training set. The inverse QSAR
problem, also de-novo design, is such that given a molecule structure representation
(usually a descriptor vector), what are the possible input structures
that satisfy the representation. This process might be considered as
a chemical design of experiments. It is clearly a non-continuous optimization problem, since
not all chemical molecules might be accessible, and cost/risk to create molecules phsically is another
critical factor. This can be made more complex, by
asking that the input structures also satisfy an experimental property
range.
\begin{itemize}
\item What methodologies can be devised to address this problem?
\item How can we ensure combinatorically created suggestions make use of the large 
chemical structure, chemical building block, and chemical reaction databases by suggesting
chemically feasible molecules (not just virtual accessible ones)? How can we improve 
synthetic accessibility predictions \cite{Boda_Seidel_Gasteiger_2007}?
\item How does a given methodology allow us to translate the
  descriptor space to a minimal region of the input space?
\item Are there internal features of a modeling algorithm (say
  hyperplanes in a SVM approach) that lets us simplify the problem?
\end{itemize}

@@

  Binary feature vectors are a common practice for chemical similarity
  searches. The typical process starts with 1. creating binary
  substructure (or other feature vectors) \cite{citeulike:8530538},
  2. creating fixed-length binary vectors of typically 1024 bits for
  reducing space requirements and speeding up further similarity
  searches (by loosing some accuracy), 3. creating further
  pre-computations for speeding up threshold based similarity searches
  \cite{doi:10.1021/ci800076s}.\\  Still, if we are interested to employ
  dynamic changes in the similarity encodings, e.g. using only a set
  of binary features, then previously done hashing or pre-computations
  might need to be redone efficiently on the fly. Finally, the major
  goal is to employ similarity searches \cite{doi:10.1021/ci200235e}
  on a scale of multiple million entries and more optimizations and
  benchmarking studies are urgently required, e.g. using GPUs
  \cite{doi:10.1021/ci1004948}, or optimizing pair-wise similarity
  calculations \cite{MINF:MINF201100050}.
%

@@

  Vectorial molecule encodings can serve as efficient approximations
  of molecules.  Sometimes non-vectorial molecular 3D shape or
  molecule kernel comparisons might be more suitabe to compare
  molecules, since they might better correlate with activities. One
  key problem is that non-vectorial encodings require to compare all
  molecules (or their 3D conformational explosions) in a pair-wise
  manner.  This becomes prohibitively expensive when considering
  millions of molecules.  Can dyadic data approaches help
  \cite{Hochreiter:2006:SVM:1159508.1159516}? Other approximations or
  cascading flows?
%
\item \emph{Using multiple annotations for improving molecular mining/predictions (chemogenomics)}
